---
title: "Day 2 EOD Activity Part 2"
format: html
author: "Jordan Sibley" 
editor: visual
execute: 
  warning: false
  message: false
  
---

## Part 2. Wild Data 

Data set acquired from https://portal.edirepository.org/nis/metadataviewer?packageid=knb-lter-and.4027.14

Aquatic Vertebrate Population Study in Mack Creek, Andrews Experimental Forest, 1987 to present

#### Looking at the data 
Description:	Vertebrate numbers and size: Annual survey of aquatic vertebrates, numbers, sizes
Number of Records:	32209
Number of Columns:	18

```{r}
# set up ----
library(tidyverse)
library(here)
library(janitor)
```

```{r}
# read in data ----
mack_verts<- read.csv(here("data", "AS00601.csv"))
```

### Data Wrangling and exploring 
```{r}
# cleaning names 
mack_verts<- clean_names(mack_verts)
```

```{r}
# practicing accessing pieces of data frame 
mc_wt_5 <- mack_verts$weight[5] 
mc_length_8_20 <- mack_verts$length1[8:20]
mc_dates <- mack_verts$sampledate
```

### Salamander subset and plots 
```{r}
# salamander subset ----
mc_salamanders<- dplyr::filter(mack_verts, species == "DITE")
head(mc_salamanders) # only has the Pacific Giant Salamanders 
```

```{r}
# scatter plot ----
# length 1 versus weight (g) in subset 

mc_salamanders_plot <- ggplot(data = mc_salamanders, 
                              aes(x = length1,
                                  y = weight))+
  geom_point(color = "darkgreen")+
  theme_minimal()+
  labs(x = "Snout-vent lenght (mm)",
       y = "Weight (g)",
       title = "Size characteristics of Dicamptodon tenebrosus",
       subtitle = "Pacific Giant Salamanders in Mack Creek, Andrews Experimental Forest 1987 to Present",
       caption = "doi:10.6073/pasta/7c78d662e847cdbe33584add8f809165")

mc_salamanders_plot
ggsave(here::here("figs", "salamander_size.png"), height = 6, width = 5)
```

### Cutthroat subset and plots 
```{r}
# make a cutthroat subset  ----

# subset with only the species ONCL
mc_trout <- dplyr::filter(mack_verts, species == "ONCL")

```

```{r}
# scatter plot ----
cutthroat_plot <- ggplot(data = mc_trout, 
                         aes(x = length1, y = weight))+
  geom_point(aes(color = reach), size = 4, show.legend = TRUE)+
  scale_color_manual(values = c("purple","pink","deepskyblue3"))+
  facet_wrap(~reach)+
  labs(x = "Snout-vent lenght (mm)",
       y = "Weight (g)",
       title = "Size characteristics of Cutthroat Trout")
  
cutthroat_plot
ggsave(here::here("figs", "cutthroat_size.png"), height = 6, width = 5)
```

